---
title: "groupProject"
author: "Audrey Huang"
date: "`r Sys.Date()`"
output: html_document
---

##Loading libraries
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(stringr)
library(ggplot2)
```

## Combining Datasets

### Read CSV Files

```{r cars}
forest_km <- read_delim("forest_area_km.csv")
forest_percent <- read_delim("forest_area_percent.csv")
global_temp <- read_delim("Global_Temperatures.csv")
```
### Checking first few columns in each csv
```{r}
head(forest_km,5) 
head(forest_percent,5) 
head(global_temp,5)
```
### Cleaning Up CSV Files
#### Convert "Country Name" header to "country"
```{r}
forest_km <- forest_km %>%
  rename(country = `Country Name`)
  head(forest_km, 5)
```
#### Convert "Country Name" header to "country"
```{r}
forest_percent <- forest_percent %>%
  rename(country = `Country Name`)
  head(forest_km, 5)
```
#### Convert "year" data type from double to char
```{r}
global_temp <- global_temp %>%
  mutate(year = as.character(year))
```

### Convert Forms from wide to long form

```{r}
countryList <- c('country', 'Country Code')

longForestKM <- forest_km %>%
   pivot_longer(!countryList, names_to = "year", values_to = "km")
longForestKM

```
```{r}

longForestPercent <- forest_percent %>%
   pivot_longer(!countryList, names_to = "year", values_to = "percent")
longForestPercent
```
### Joining Table Together for Forest
```{r}
forest_km_percent <- left_join(longForestKM, longForestPercent)
forest_km_percent
```
### Inner Joining global temp together to create a final set
```{r}
final_set2 <- inner_join(global_temp, forest_km_percent)
final_set2
```

### Q: At which average temperature do the forests in that area tend to thrive?
```{r}
final_set2 %>% filter(year==2020 | year==1995) %>% group_by(year) %>% 
  ggplot(aes(x=average_temperature, y=percent, col=year)) + geom_point() + 
  geom_smooth(method=lm, se=FALSE)
```

From this plot, we can see that in both 1995 and 2020 (the earliest and latest 
recorded years in the dataset), there is a positive correlation between average 
temperature and percent of forestry in the country. 


But at which temperature do the forests tend to thrive? 

```{r}
final_set2 %>% filter(year==2020 | year==1995) %>% group_by(year) %>% 
  filter(rank(-percent)==1) %>% summarize(country, average_temperature, percent)
```

The temperature at which the highest percentage of forestry occurs at around 79 degrees. 


But does this mean that forestry always thrives at 79 degrees?

```{r}
final_set2 %>% 
  group_by(temperature_range = factor(round(average_temperature/10)*10)) %>% 
  reframe(average_percent=mean(percent, na.rm=TRUE)) %>% 
  arrange(rank(-average_percent))
```

Actually, the highest average percentage of forestry occurred in the 35-44 
degree range, with 38% forestry. The range holding the instance of highest 
percentage of forestry (which was 79 degrees) is the next ranked temperature 
range, with 35% forestry on average across the countries. 


What if "thrive" was taken to mean growth?

```{r}
final_set2 %>% filter(year==1995 | year==2020) %>% group_by(country) %>% 
  reframe(average_temp=mean(average_temperature), growth=abs(diff(percent))) %>% 
  ggplot(aes(x=average_temp, y=growth)) + geom_point() + 
  geom_smooth(method=lm, se=FALSE)
```

From this graph, we can see that there is a clear correlation between average 
temperature and growth (difference in percent forestry in 2020 and 1995). 


What is the temperature with the most growth?

```{r}
final_set2 %>% filter(year==1995 | year==2020) %>% group_by(country) %>% 
  reframe(average_temp=mean(average_temperature), growth=abs(diff(percent))) %>% 
  filter(rank(-growth)==1) %>% summarize(country, average_temp, growth)
```

The largest growth in percent forestry was 20% in Nicaragua, at an average 
temperature of 81 degrees. 


Which temperature range has the most growth on average?

```{r}
final_set2 %>% filter(year==1995 | year==2020) %>% group_by(country) %>% 
  reframe(average_temp=mean(average_temperature), growth=abs(diff(percent))) %>% 
  ungroup() %>% 
  group_by(temperature_range=factor(round(average_temp/10)*10)) %>% 
  reframe(average_growth = mean(growth, na.rm=TRUE)) %>% 
  arrange(rank(-average_growth))
```

The 75-84 degree temperature range shows the greatest average growth. This is 
consistent with the results from the previous section. 

